---
title: "ANTH 241 Homework 3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Anth241Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval=F}
# Always start your homework by copying and pasting the line below
library(Anth241Hmwk)
```
In this assignment you will be looking at anthropometric variation.
For each question you will run code to produce plots and then answer
some follow-up questions. **Make sure** to read all the way to the end
of each question to answer the follow-up questions.

## 1. Growth trajectory

### Part A

Run the below code and copy/paste the resulting plot into your document.
Note: The plot you are doing is called a "Preece-Baines" growth curve.
This is a "distance curve" in that it is showing "how far" the kid has
gotten in their stature ("standing height").

```{r eval=FALSE}
Fig11_07()
```

### Part B

What is the approximate age of the adolescent growth spurt on this plot?
How do you know? (Hint: it should be the age at which growth starts to
accelerate).

## 2. Growth Velocity

### Part A

Now we will get the velocity curve, or in other words the speed at which
the kid is growing at any given age.  If you like calculus (most don't),
then the velocity curve is the first derivative of the distance curve.
Copy/paste the plot from this code to your document.

```{r eval=FALSE}
# warning! This will only work if you have already run Fig11_07
Fig11_08() 
```

### Part B
If you had two velocity curves, one for boys and one for girls, what is
the primary way in which you would expect them to differ?  Recall when
you were in grade school how girls generally got their adolescent growth
spurt earlier than boys.

## 3. Growth Acceleration

### Part A

Paste the resulting plot into your document.

```{r eval=FALSE}
Fig11_08b()
```

### Part B

Explain how the timing of the adolescent growth spurt is related to
adult height.

## 4. Allen’s rule

### Part A

Run Fig11_14() and paste the resulting plot into your document.

```{r eval=FALSE}
Fig11_14()
```

### Part B

How do these plots represent Allen’s Rule?

### Part C
Use Allen’s rule to explain how the environment of the first rabbit
likely differs from that of the second rabbit:

```{r echo=FALSE, out.width = "70%", fig.align = "center"}
knitr::include_graphics(file.path(here::here(),"images","Rabbit1.png"))

```


```{r echo=FALSE, out.width = "70%", fig.align = "center"}
knitr::include_graphics(file.path(here::here(),"images","Rabbit2.png"))

```

