---
title: "ANTH 241 Homework 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Anth241Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval=F}
# Always start your homework by copying and pasting the line below
library(Anth241Hmwk)
```

## 1. Hardy-Weinberg equilibrium for ABO blood group

The script below asks for the number of individuals with type A, B, AB,
and O blood. Only for AB and O phenotypes are the genotypes known (AB and
OO,respectively). Type A includes AA and AO while type B includes BB and
BO. The script uses something called the expectation maximization
algorithm to find the allele frequencies (p [frequency of A allele], q
[frequency of B allele], and r [frequency of O allele]).  You will be
asked for phenotype counts and there are suggested phenotype counts.  Use
the suggested counts.

The output will be "Obs" and "Exp" which are the observed phenotype
counts and the expected counts if the locus really was in Hardy-Weinberg
equilibrium.  You will also get $N (the total number of individuals), $p
(the estimated A allele frequency), $q (the estimated B allele
frequency), and $r (the estimated O allele frequency).

### Part A
```{r eval=FALSE}
# run entering the suggested phenotype counts
# and copy output (Obs & Exp and $N, $p, $q, and $r) and paste
# to your document
P_56()
```

### Part B
This time you are going to enter your own counts
```{r eval=FALSE}
# run entering whatever phenotype counts you want
# and copy output (Obs & Exp) and paste
# to your document
sto=P_56()
```

```{r eval=FALSE}
# This time you did not see the $N, $p, $q, and $r. Why is
# that? It is because they got stored in "sto." To see them
# copy and paste the line below to your console:
sto
```

### Part C

Now for some real fun!  The reason that we did that storage to "sto"
(my favorite name for storage) is so that we can use the stuff we
stored there to find the expected count of AB blood types, which is
$N \times 2 \times p \times q$.  But 
["Houston, we have a problem"](https://en.wikipedia.org/wiki/Houston,_we_have_a_problem)
There is no $\times$ symbol on the keyboard, nor can we type "p" or
"q" thinking that the keyboard will recognize those as allele
frequencies.  Any programming language or scripting language that
you look at will use * (the asterisk symbol, which is on the keyboard)
in place of the $\times$ symbol. So, the code below will gives us
what we want.

```{r eval=FALSE}
# Paste this output to your document.  It should be
# the same as from the "expected" value for the AB phenotype
sto$N*2*sto$p*sto$q

```

## 2. Hardy-Weinberg and probability values

In the next problem we are going to look at genotype counts for a
bi-allelic locus (so AA, AB, and BB genotypes) to see if the genotypes
are in Hardy-Weinberg equilibrium.  This script will take the observed
genotype counts and return the expected genotype counts if the genotypes
were in H-W equilibrium.  But wait, there's more.  The script will also
return the probability value ("p-value") that the genotypes are in
Hardy-Weinberg equilibrium.  This is found by finding the probability
of getting the observed AA genotype count under H-W equilibrium or
any more extreme count (but will under H-W equilibrium).  Aside from
just getting the information in your console, you will also get a
graphical representation of how the probability value is found.

### Part A

```{r eval=FALSE}
# Run this and copy/paste console output and graph to your document
Tab3_2()
```

### Part B

OK, that was running the default values, but now we are going to
run an example where we start in H-W equilibrium (so p-value=1) and
move further and further away from H-W equilibrium by reducing N from
100 to lower values for the BB genotype count. For each run copy
the console output to your document but **do not** copy and paste
the graphs (that would be a lot of graphs!).

```{r eval=FALSE}
# Copy/paste the two lines below to setup the allele frequencies
p=1/3
q=1-p # remember, in bi-allelic system q=1-p
```


```{r eval=FALSE}
# Run the line below and copy/paste output to document
Tab3_2(100*p^2,100*2*p*q,100*q^2) # This is in H-W equil, so p=1
```

```{r eval=FALSE}
# Run the line below and copy/paste output to document
Tab3_2(100*p^2,100*2*p*q,50*q^2)
```

```{r eval=FALSE}
# Run the line below and copy/paste output to document
Tab3_2(100*p^2,100*2*p*q,10*q^2)
```
### Part C

Which of the three runs above gave you the lowest probability value?
(i.e., which of the runs was furthest from being in H-W equilibrium?)
Note that this is due to dropping the number of total individuals for
the BB genotype.

## 3. Balancing/Stabilizing selection

We will start with the default setting which is for "balancing"
selection.  In other words, the heterozygote has the highest relative
fitness (=1.0), while the homozygotes have fitness values that are
greater than 0.0 and less than 1.0.  In this case w=c(0.7635,1.0,0.0001)
where the \emph{w} values are the relative fitnesses.  In "R-ese" the
$c(\cdot)$ is the "combine" function where the list with commas gets
combined into a vector, so the relative fitness for the AA genotype is
0.7635, for the AS genotype (sickle-cell trait) is 1.0, and for the
SS genotype (sickle-cell disease) is 0.0001.

### Part A

```{r eval=FALSE}
# Run and copy/paste the graph to your document
Fig3_9(p.start=runif(100))
```

### Part B

What does each line on the graph represent?

## 4. Heterozygote disadvantage

Now we will look at homozygote advantage, where the relative fitness
of the AA and BB genotypes (the homozygotes) are  1.0 and the
heterozygotes have a relative fitness greater than 0.0 but less than
1.0.

### Part A

```{r eval=FALSE}
# Run me and copy/paste the plot to your document
Fig3.9(c(1,.6,1),runif(100))
```

### Part B

Look up ["fixed allele"](https://en.wikipedia.org/wiki/Fixed_allele) at
Wikipedia and explain how this relates to what you see in the graph
(with the realization that either allele, "A" or "B", can go to
fixation).

## 5. Adaptive topography

### Part A
In this section you will do plots of average relative fitness for the
sickle-cell example and heterozygote disadvantage and paste these graphs
to your homework.

```{r eval=FALSE}
# Balanced polymorphism
Fig3_13()
```

```{r eval=FALSE}
# Heterozygote disadvantage
Fig3_13(c(1,.6,1),c(0,1),c(0,1),tan=.05)
```

### Part B

In the second plot, if you drew arrows to indicate the direction of
evolution, would they point toward or away from the middle of the plot? 
Remember that evolution by selection is about "hill-climbing" in the
adaptive topography.  To wit, selection cannot take you "down hill".

## 6. Genetic drift

**Note: there is nothing to turn in from this section**

Genetic drift is one of the harder concepts to grasp, so it will behoove
you to look at a number of graphs. The scripts Fig3_17 and Fig3_20 were
written in part so that people could “play around” and look at a bunch
of graphs. You will see graphs like those out of Fig3_17 on your exam 
and will be asked to sort them by increasing sub-population size.
Similarly, you will see graphs like Fig3_20 but with generations out
of order and you will be asked to sort them by generation.

#### Stochastic simulation of genetic drift

```{r eval=FALSE}
Fig3_17(10^0,jit=0.1) # i.e., deme sizes are all = 1
```

```{r eval=FALSE}
Fig3_17(10^1) # i.e., deme sizes are all = 10
```

```{r eval=FALSE}
Fig3_17(10^2) # i.e., deme sizes are all = 100
```

```{r eval=FALSE}
Fig3_17(10^3) # i.e., deme sizes are all = 1,000
```


```{r eval=FALSE}
Fig3_17(10^4) # i.e., deme sizes are all = 10,000
```

#### Deterministic simulation of genetic drift

The default is that there are five individuals per deme, so 
$2 \times 5=10$ alleles.  You will be asked how many allele at
the dawn of time are A alleles.  Simplest is to answer 5.

```{r eval=FALSE}
Fig3_20() # and answer 5 for number of A alleles
```

```{r eval=FALSE}
# This one will "drift slower", but you would still be able to
# order the graphs
Fig3_20(5) # and answer 10 for number of A alleles
```
