---
title: "ANTH 241 Homework 2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Anth241Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval=F}
# Always start your homework by copying and pasting the line below
library(Anth241Hmwk)
```

In this assignment you will be examining how different population 
parameters affect coalescence.  A broad hint: coalescence is another
way of looking at genetic drift.  Small demes experience genetic drift
more than large demes.  Similarly, small demes coalesce more quickly
(i.e., reach the most recent common ancestor at a smaller tree depth)
while larger demes coalesce more slowly (i.e., reach the most recent
common ancestor at a greater tree depth).  You will also notice that
coalescent events become "slower" (more rare) as you work up the tree
following previous coalescent events.

## 1. Size in static populations

### Part A

The following code will simulate two coalescent trees drawn on the
same time scale.  Remember that deeper trees represent larger demes.

```{r eval=FALSE}
Fig9_6()

```

### Part B
If you didn’t know which population was larger (right or left side
plot), how could you figure this out from the plots that you just made?
Using this device, which population was larger?  The one on the left or 
the one on the right?

### Part C
If you changed the size of the second population to 10,000 instead of
its default value of 1,000, what would you expect to happen to the plot?
If you are uncertain, run the following code:

```{r eval=FALSE}
Fig9_6(100,10000)
```

## 2. Mismatch distributions

The mismatch distribution is the distribution (well, that was redundant)
of mutational differences between all pairs of individuals.  We can 
simulate that by placing random mutations on the coalescent tree. 
Because we assume a constant mutation rate and that all mutations are
unique event polymorphisms, longer branches "see" more mutations.

### Part A
Run the Fig9_7 script and copy the tree and the mismatch distribution to
your document.  Be sure that you copy both the tree and the mismatch 
distribution to your homework.


Warning: if you are using RStudio, the mismatch distribution will appear
in a separate window and the tree will be in the lower right panel. You 
may have to hunt around a bit for the window with the mismatch
distribution.  On a Windows machine it may appear in the bottom ribbon.
For a Mac I am not sure where that new window will appear. If you are
using R, both plots will appear in a separate window with the mismatch
distribution blocking the tree. Just drag the histogram window out of
the way.

```{r eval=FALSE}
Fig9_7()
```

### Part B

What is being compared to produce the mismatch distribution?

### Part C

What aspect of the tree causes the mismatch distribution to appear 
“ragged”?

## 3. Population expansion

### Part A
Run the script for a static (non-increasing) population: Fig9_12(1000,0)
and for an expanding population: Fig9_12() and copy the trees and
mismatch distributions to your document. Your run with the defaults
(i.e., Fig9_12()) is doing an expanding population.

```{r eval=FALSE}
# Static population
Fig9_12(1000,0)
```

Be sure to copy the tree and the mismatch distribution and paste to your
document.  Be sure to close any open windows after this so that things
do not get confusing.

```{r eval=FALSE}
# Expanding population
Fig9_12()
```

### Part B
Explain why the mismatch distribution for the expanding population no
longer appears ragged.

### Part C
Explain how you could use the shape of the mismatch distribution to
figure out if a population has experienced an expansion event in the
past.  Note: you will see mismatch distributions and coalescent trees
on you exam and be asked which ones are from static populations and
which ones are from expanding populations.
